```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r}
library(tidyverse)
library(qqman)
library(stringr)
```


# Introduction
You've recently discovered that there are in fact two types of cholesterol-- both good (HDL) and bad (LDL). You are worried that you may have a genetic predisposition to having high levels of bad cholesterol and decide to investigate what genes may be associated with bad cholesterol. Here, we'll do a basic exploration of the data before investigating the specific genetic effects in subsequent lectures. 

# Problem 1
**Access the LDL summary statistics from the GLGC Consortium's genome-wide association study. Visualize the summary statistics both as a Manhattan Plot and as a Q-Q plot. What chromosome(s) appear to have genome-wide significant hits?** <br>

_Hint: use the _**qqman** _package in **R** to visualize the data, which can be accessed through_ **LDHub**. 

```{r}
pathToDataFile <- "./jointGwasMc_LDL.txt.gz"
sumstats <- as.data.frame(data.table::fread(paste0("\"C:\\Program Files\\Git\\bin\\sh.exe\" zcat < ", pathToDataFile), showProgress = FALSE))
```

```{r}
sumstats <- sumstats %>% mutate(snp = gsub("chr", "", sumstats$SNP_hg19))
sumstats <- sumstats %>% separate(snp, sep = ":", into = c("CHR", "BP")) %>%
  filter(`P-value` != 0) %>% na.omit()
sumstats$CHR <- as.numeric(sumstats$CHR)
sumstats$BP <- as.numeric(sumstats$BP)
manhattan(sumstats, snp = "rsid", p = "P-value")
```


\newpage

# Problem 2
**As part of the GLGC Consortium, the group analyzed data for a different SNP array, the Metabochip. Visualize the summary statistics both as a Manhattan Plot and as a Q-Q plot. What chromosome(s) appear to have genome-wide significant hits?**<br>

_Hint: use the Metabochip summary statistics can be found on the same page as the GWAS summary statistics_



\newpage

# Problem 3
**Compute the measure of systematic inflation ($\lambda_{GC}$) associatd with the summary statistics in Problem 1 and Problem 2. For which SNP array are the summary statistics more inflated? **

_Hint:_  $\lambda_{GC}$ = median($\chi^2$) `/ 0.4549364` _where the last number comes from_ `qchisq(0.5,1)`

\newpage
